Expected Dependency Pair Match:\\ Predicting {HTER} improvements with expected syntactic structure

### 2009-05-11 mtjournal.tex
whitespace and obsolete comment cleanup
basic structure of hter1 section in place

%
In this work, we use a state-of-the-art PCFG (the first stage of
\inlinecite{charniak-johnson:2005:ACL}) and context-free head-finding rules
\cite{magerman95headfinding} to generate a 50-best list of dependency
trees for each hypothesis and reference translation.  We use the
parser's default (English-language) Wall Street Journal training
parameters.  Head-finding uses the Charniak parser's rules, with three
modifications: prepositional and complementizer phrases choose nominal
and verbal heads respectively (rather than functional heads) and
auxiliary verbs are modifiers of main verbs (rather than the
converse).

Having constructed the dependency tree, we label the arcs as $d
\stackrel{A/B}{\to} h$, where the arc label $A/B$ between dependent
$d$ and its head $h$ is composed of $A$ (the lowest constituent-label
headed by $h$ and dominating $d$) and $B$ (the highest constituent
label headed by $d$). 
\begin{figure}
  % \centering
  \Tree
  [.\headlabel{S}{stumbled}
    [.\headlabel{NP}{cat}
      [.\headlabel{DT}{the} the ]
      [.\headlabel{NN}{cat} cat ]
    ]
    [.\headlabel{VP}{stumbled} 
      [.\headlabel{VBD}{stumbled} stumbled ]
    ]
  ]
  \\
  \begin{center}
    \includegraphics[scale=0.6]{dpm-example-depextract.eps}
  \end{center}
  \caption{An example constituent tree (heads of each constituent are
    listed small below the label) and the labelled dependency tree
    derived from it using the strategy described in
    section~\ref{sec:paradigm}.}
  \label{fig:depextract}
\end{figure}
For example, in figure~\ref{fig:depextract}, the S node is the
lowest node headed by \emph{stumbled} that dominates \emph{cat}, and
the NP node is the highest constituent label headed by \emph{cat}, so
the arc between \emph{cat} and \emph{stumbled} is labelled \arclabel{S/NP}.
%
This strategy is very similar to one adopted in the reference
implementation of labelled-dependency \textsc{SParseval}, and may be
considered as an approximation of the rich semantics generated by LFG
parsers \cite{cahill04lfg} or another heavily knowledge-engineered
parser, but with much less knowledge-engineering required.

@@

For some of the experiments presented here (sections~\ref{sec:hter1}
and~\ref{sec:hter2}), we have multiple translations of the same source
segment, each from a different translation engine and each with its
own HTER score.  For these translations, reporting correlations of
metric $m$ with absolute HTER scores can conflate $m$'s power in
identifying which of two candidate translations is better with $m$'s
(and HTER)'s ability to distinguish which source segments are
more difficult to translate.

To avoid this conflation, on the HTER corpora we report correlations of
the mean-removed $m$ and HTER scores. Mean-removal  lowers the
correlation $r$ values reported but
ensures that the reported correlations are among differences in the
translations rather than among differences in the underlying segments.

@@

\section{Per-document correlations with HTER}
\label{sec:hter1}
TO DO: INTRODUCE THIS SECTION

@@

The GALE 2.5 translation corpus is made up of system translations into
English from three sites.
% MO added:
The three sites all use system combination to integrate results from
multiple systems, some of which are phrase-based and some which may
use syntax on either the source or target side. No system provided
system-generated parses.
%
The corpus being translated comes from
Arabic and Chinese in four genres: \texttt{bc} (broadcast
conversation), \texttt{bn} (broadcast news), \texttt{nw} (newswire),
and \texttt{wb} (web text), with corpus sizes shown in
table~\ref{tab:galestats}.
\begin{table}
%  \centering
  \begin{tabular}{r|rr|rr|rr}
     & \multicolumn{2}{c|}{Arabic} & \multicolumn{2}{c|}{Chinese}
     & \multicolumn{2}{c}{Total}\\
     & doc & sent & doc & sent & doc   & sent\\
     \hline
     \texttt{bc}    & 59  & 750 & 56 & 1061 & 115 & 1811\\
     \texttt{bn}    & 63  & 666 & 63 & 620  & 126 & 1286\\
     \texttt{nw}    & 68  & 494 & 70 & 440  & 138 & 934 \\
     \texttt{wb}    & 69  & 683 & 68 & 588  & 137 & 1271\\
     \hline
     Total & 259 & 2593& 257& 2709 & 516 & 5302\\
  \end{tabular}
  \caption{Corpus statistics for the GALE 2.5 translation
    corpus.}
  \label{tab:galestats}
\end{table}
The corpus includes one English reference translation $r_i$
\cite{gale08phase2_5references} for each sentence $i$ and a system
translation $t_{i,z}$ for each of the three systems
$z$. Additionally, each of the system translations of each segment $i$
has a corresponding human-targeted reference aligned at the sentence
level, so we have available the HTER score of each segment
$s_{\textsc{hter}}(t_{i,z})$ at both the sentence and document level.

  prediction:
    Mean-subtracted HTER per-document

  base measure EDPM: F[1g,2g,dl,lh],n=50,gamma=0.25

  [EXPT]
\begin{verbatim}
            Ar     Zh    All
   TER     0.51   0.32  0.44
   BLEU_4 -0.42  -0.33 -0.37
   EDPM   -0.60* -0.39 -0.50*
\end{verbatim}
  point: base EDPM does better on all docs, and on Arabic. difference
  is not quite significant at the p=0.05 on the Chinese documents

  Note 1: also tried pairwise deltas per-document, with similar
  results, as presented in the MetricsMATR competition submission.

  Note 2: when using per-segment mean-subtracted, r values are smaller
  and *Arabic* was the language where EDPM did not perform best.

\begin{verbatim}
               Ar     Zh     All
      TER     0.38*  0.04   0.19
      BLEU_4 -0.10  -0.16  -0.14
      EDPM   -0.31  -0.19* -0.24*
\end{verbatim}

  include genre tables here as well -- cut out again if no story.

>>>>

In this work, we use a state-of-the-art PCFG (the first stage of
\inlinecite{charniak-johnson:2005:ACL}) and context-free head-finding rules
\cite{magerman95headfinding} to generate a 50-best list of dependency
trees for each hypothesis and reference translation.  We use the
parser's default (English-language) Wall Street Journal training
parameters.  Head-finding uses the Charniak parser's rules, with three
modifications: prepositional and complementizer phrases choose nominal
and verbal heads respectively (rather than functional heads) and
auxiliary verbs are modifiers of main verbs (rather than the
converse).

Having constructed the dependency tree, we label the arcs as $d
\stackrel{A/B}{\to} h$, where the arc label $A/B$ between dependent
$d$ and its head $h$ is composed of $A$ (the lowest constituent-label
headed by $h$ and dominating $d$) and $B$ (the highest constituent
label headed by $d$). 
\begin{figure}
  \Tree
  [.\headlabel{S}{stumbled}
    [.\headlabel{NP}{cat}
      [.\headlabel{DT}{the} the ]
      [.\headlabel{NN}{cat} cat ]
    ]
    [.\headlabel{VP}{stumbled} 
      [.\headlabel{VBD}{stumbled} stumbled ]
    ]
  ]
  \\
  \begin{center}
    \includegraphics[scale=0.6]{dpm-example-depextract.eps}
  \end{center}
  \caption{An example constituent tree (heads of each constituent are
    listed small below the label) and the labelled dependency tree
    derived from it using the strategy described in
    section~\ref{sec:paradigm}.}
  \label{fig:depextract}
\end{figure}
For example, in figure~\ref{fig:depextract}, the S node is the
lowest node headed by \emph{stumbled} that dominates \emph{cat}, and
the NP node is the highest constituent label headed by \emph{cat}, so
the arc between \emph{cat} and \emph{stumbled} is labelled \arclabel{S/NP}.
%
This strategy is very similar to one adopted in the reference
implementation of labelled-dependency \textsc{SParseval}, and may be
considered as an approximation of the rich semantics generated by LFG
parsers \cite{cahill04lfg} or another heavily knowledge-engineered
parser, but with much less knowledge-engineering required.

@@

For some of the experiments presented here (sections~\ref{sec:hter1}
and~\ref{sec:hter2}), we have multiple translations of the same source
segment, each from a different translation engine and each with its
own HTER score.  For these translations, reporting correlations of
metric $m$ with absolute HTER scores can conflate $m$'s power in
identifying which of two candidate translations is better with $m$'s
(and HTER's) ability to distinguish which source segments are
more difficult to translate.

To avoid this conflation, on the HTER corpora we report correlations
on $\bar{\mu}(m)$
and $\bar{\mu}(\textrm{HTER})$ (rather than measure $m$ and HTER), where
$\bar{\mu}$ represents mean-removal:
\begin{equation}
  \label{eq:meansub}
  \bar{\mu}(m(t_i)) = m(t_i) - \sum_{j=1}^I\frac{m(t_j)}{I}
\end{equation}
Mean-removal lowers the correlation $r$ values reported but ensures
that the reported correlations are among differences in the
translations rather than among differences in the underlying segments.

@@

\section{Per-document correlations with HTER}
\label{sec:hter1}
In this section, we move from an exploration of the space of possible
DPM metrics to a single chosen DPM-derived metric (EDPM).  Here, we
compare EDPM to other metrics in correlation with document-level HTER
performance.

@@

The GALE 2.5 translation corpus is made up of system translations into
English from three sites.
% MO added:
The three sites all use system combination to integrate results from
multiple systems, some of which are phrase-based and some which may
use syntax on either the source or target side. No system provided
system-generated parses.
%
The corpus being translated comes from
Arabic and Chinese in four genres: \texttt{bc} (broadcast
conversation), \texttt{bn} (broadcast news), \texttt{nw} (newswire),
and \texttt{wb} (web text), with corpus sizes shown in
table~\ref{tab:galestats}.
\begin{table}
  \begin{tabular}{r|rr|rr|rr}
     & \multicolumn{2}{c|}{Arabic} & \multicolumn{2}{c|}{Chinese}
     & \multicolumn{2}{c}{Total}\\
     & doc & sent & doc & sent & doc   & sent\\
     \hline
     \texttt{bc}    & 59  & 750 & 56 & 1061 & 115 & 1811\\
     \texttt{bn}    & 63  & 666 & 63 & 620  & 126 & 1286\\
     \texttt{nw}    & 68  & 494 & 70 & 440  & 138 & 934 \\
     \texttt{wb}    & 69  & 683 & 68 & 588  & 137 & 1271\\
     \hline
     Total & 259 & 2593& 257& 2709 & 516 & 5302\\
  \end{tabular}
  \caption{Corpus statistics for the GALE 2.5 translation
    corpus.}
  \label{tab:galestats}
\end{table}
The corpus includes one English reference translation $r_i$
\cite{gale08phase2_5references} for each sentence $i$ and a system
translation $t_{i,z}$ for each of the three systems $z$. Additionally,
each of the system translations of each segment $i$ has a
corresponding human-targeted reference aligned at the sentence level,
so we have available the HTER score of each segment
$s_{\textrm{HTER}}(t_{i,z})$ at both the sentence and document level.

As discussed in section~\ref{sec:paradigm}, we report correlations
here between $\bar{\mu}(m)$ and $\bar{\mu}(\textrm{HTER})$, rather than
$m$ and HTER directly, to abstract away from variations in the
underlying documents.

In table~\ref{tab:hterperdoc}, we show per-document Pearson's $r$
between $\bar{\mu}$EDPM and $\bar{\mu}$HTER, as well as two other
baselines: $\bar{\mu}$TER and $\bar{\mu}$BLEU$_4$.  
\begin{table}
  \begin{tabular}{r|rrrr|rr|r}
    \hline
    $r$ vs. $\bar{\mu}$HTER & \multicolumn{1}{c}{\texttt{bc}}
    & \multicolumn{1}{c}{\texttt{bn}} &
    \multicolumn{1}{c}{\texttt{nw}} & \multicolumn{1}{c}{\texttt{wb}}
    & \multicolumn{1}{|c}{all Arabic} & \multicolumn{1}{c|}{all Chinese}
    & \multicolumn{1}{c}{all} \\
    \hline
    $\bar{\mu}$TER
    &  0.59 &  \textbf{0.35} &  \textbf{0.47}&  \textit{0.17}
    & \textbf{0.54} & \textbf{0.32} &  0.44\\
    $\bar{\mu}$BLEU
    & -0.42 & \textbf{-0.32} & \textbf{-0.46}&  \textbf{-0.27}
    & -0.42  &  \textbf{-0.33} & -0.37 \\
    $\bar{\mu}$EDPM
    & \textbf{-0.69} & \textbf{-0.39} & \textbf{-0.47} &
    \textbf{-0.27}
    & \textbf{-0.60} & \textbf{-0.39} & \textbf{-0.50} \\
    \hline
  \end{tabular}
  \caption{Per-document correlations of $\bar{\mu}$EDPM and others to
    $\bar{\mu}$HTER, by genre and by source language. Bold numbers are
  within 95\% significance of the best per column; italics indicate
  that the sign of the $r$ value has less than 95\% confidence.}
  \label{tab:hterperdoc}
\end{table}
We see that the EDPM has the highest correlation in each of the
subcorpora created by dividing by genre or by source language, as well
as the corpus as a whole.  Not every difference is above 95\%
confidence, but the trend is the same across all of the
subcorpora. EDPM's strengths are particularly clear in the
unstructured domains (\texttt{wb} and \texttt{bc}).

TO DO: NOTE RE: DELTAS

TO DO: NOTE RE: SEGMENT-LEVEL CORRS

%   Note 2: when using per-segment mean-subtracted, r values are smaller
%   and *Arabic* was the language where EDPM did not perform best.
% \begin{verbatim}
%                Ar     Zh     All
%       TER     0.38*  0.04   0.19
%       BLEU_4 -0.10  -0.16  -0.14
%       EDPM   -0.31  -0.19* -0.24*
% \end{verbatim}
